Implementation of a lattice Boltzmann method for immiscible two-phase flow simulations using the level set method

### 2015-11-14 Paper/paper.tex

As boundary conditions and initial conditions we use,
\begin{align}
  \vec u_i(x,t) &= 0 \text{, on } \delta\Omega_i \backslash \Gamma \text{,} \\
  \vec u_i(x,0) &= \vec u_{ini,i}(x) \text{, in } \Omega_i \text{.}
\end{align}
At the interface $\Gamma$ we have the influence of the surface tension and a no-slip condition, i.e. continuity of the velocities between the fluid phases. Therefore, we get the jump conditions,
\begin{align}
  [\vec u] &= 0 \text{,  } \\
  [2\mu S - p \mathbf{I}] \cdot \vec n &= 2 \sigma \kappa \vec n \text{.}
	\label{jumpconditions}
\end{align}
Here, S is the shear rate tensor $S=\frac12 \left( \nabla \vec u + \nabla {\vec u}^T\right)$, $\sigma$ the surface tension, $\mu = \varrho \nu$ the dynamic viscosity and $\kappa$ the curvature of the interface with respect to its outer normal $\vec n$. Brackets denote the jump of a quantity at the interface, 
\begin{equation}
	[q](\vec x) = \lim_{\epsilon \rightarrow 0} (q(\vec x - \epsilon \vec n) - (q(\vec x + \epsilon \vec n)) \text{ with } \vec x \in \Gamma. % I also turned this around, according to [p].
\end{equation}
In \cref{intro_phases} this is the jump from $\Omega_1$ to $\Omega_2$. For the other direction the sign changes.

>>>>

As boundary conditions and initial conditions we use,
\begin{align}
  \vec u_i(x,t) &= 0 \text{, on } \delta\Omega_i \backslash \Gamma \text{,} \\
  \vec u_i(x,0) &= \vec u_{ini,i}(x) \text{, in } \Omega_i \text{.}
\end{align}
At the interface $\Gamma$ we have the influence of the surface tension and a no-slip condition, i.e. continuity of the velocities between the fluid phases. Therefore, we get the jump conditions,
\begin{align}
  [\vec u] &= 0 \text{,  } \\
  [2\mu S - p \mathbf{I}] \cdot \vec n &= 2 \sigma \kappa \vec n \text{.}
	\label{jumpconditions}
\end{align}
Here, S is the shear rate tensor $S=\frac12 \left( \nabla \vec u + \nabla {\vec u}^T\right)$, $\sigma$ the surface tension, $\mu = \varrho \nu$ the dynamic viscosity and $\kappa$ the curvature of the interface with respect to its outer normal $\vec n$. Brackets denote the jump of a quantity at the interface, 
\begin{equation}
	[q](\vec x) = \lim_{\epsilon \rightarrow 0} (q(\vec x - \epsilon \vec n) - (q(\vec x + \epsilon \vec n)) \text{ with } \vec x \in \Gamma. % I also turned this around, according to [p].
\end{equation}
In Fig. \ref{intro_phases} this is the jump from $\Omega_1$ to $\Omega_2$. For the other direction the sign changes.

There are different approaches to solve the immiscible two-phase problem. The most spread among these approaches is the colour gradient method of Gunstensen and Rothman \cite{GRcg1,GRcg2,GRcg3}. The downside of this method is that it is only applicable for small density and viscosity differences. Another approach is the concept of {\it interaction potentials} by Shan and Chen \cite{ShanChen1,ShanChen2}. This method actually models miscible fluids and can only approximately describe the behavior of immiscible fluids. Furthermore there are free surface methods which only take one fluid in account for computation (\cite{Koerner}). They work best for big density and viscosity ratios.
%TODO: Mehr Referenzen, siehe Mail

In this publication we use a hybrid lattice Boltzmann level set method. Thereby, the interface motion is calculated by the level set method. The input data for the level set method is given by the lattice Boltzmann method, which is calculated independently for each fluid. This method has the advantage that it allows big ranges of density and viscosity ratios between the fluid phases 
and has a high accuracy but requires only little memory %TODO das ist zu unscharf, glaube ich
.
In \cref{intro_phases} this is the jump from $\Omega_1$ to $\Omega_2$. For the other direction the sign changes.
